Applications of Many-Core Technologies to On-line Event Reconstruction in High Energy Physics Experiments

### 2013-11-18 papers/2013/NSS/NSS.tex
add comment about opencl failing on nvidia

\subsection{Data processing in \textsc{opencl}}
We also measure the track fitting algorithm latency in an
implementation using \textsc{opencl}. The \textsc{opencl} tools have
the advantage of being an open standard, while the \textsc{cuda} tools
are only compatible with Nvidia GPUs. Additionally, \textsc{opencl}
can work also on CPUs and on the Xeon Phi. For our tests, due to
limitations on the available space for storing data on the GPU, we
only test the algorithm on up to 1000 events (\textit{i.e.} about 2
million combinations to fit). In Fig.~\ref{fig:opencl_timing} we show
the results of algorithm latency measurements for three different
modes: running the serial algorithm on the CPU (single-core), running
the \textsc{opencl} algorithm on the CPU (multi-core), and running the
\textsc{opencl} algorithm on the AMD Radeon GPU. The \textsc{opencl}
implementation of the algorithm on the CPU provides a significant
speedup---about a factor of 5---over running the algorithm serially on
the CPU. Running the same \textsc{opencl} algorithm on the GPU
provides an even greater speedup due to the increased number of cores
and parallel threads that can be run, though there is additional
latency to copy data into and out of the GPU that makes running on the
GPU take longer for small numbers of combinations to fit.
\begin{figure}[!t]
  \centering
  \includegraphics[width=0.9\linewidth]{figures/TotalTimeOpencl_1000}
  \caption{The timing of the \textsc{opencl} algorithm as a function of the
    number of track fits. We compare timing running the algorithm serially on a CPU, 
    using \textsc{opencl} on the CPU, and on an AMD GPU.}
  \label{fig:opencl_timing}
\end{figure}

>>>>

\subsection{Data processing in \textsc{opencl}}
We also measure the track fitting algorithm latency in an
implementation using \textsc{opencl}. The \textsc{opencl} tools have
the advantage of being an open standard, while the \textsc{cuda} tools
are only compatible with Nvidia GPUs. Additionally, \textsc{opencl}
can work also on CPUs and on the Xeon Phi. For our tests, due to
limitations on the available space for storing data on the GPU, we
only test the algorithm on up to 1000 events (\textit{i.e.} about 2
million combinations to fit). In Fig.~\ref{fig:opencl_timing} we show
the results of algorithm latency measurements for three different
modes: running the serial algorithm on the CPU (single-core), running
the \textsc{opencl} algorithm on the CPU (multi-core), and running the
\textsc{opencl} algorithm on the AMD Radeon GPU. The \textsc{opencl}
implementation of the algorithm on the CPU provides a significant
speedup---about a factor of 5---over running the algorithm serially on
the CPU. Running the same \textsc{opencl} algorithm on the GPU
provides an even greater speedup due to the increased number of cores
and parallel threads that can be run, though there is additional
latency to copy data into and out of the GPU that makes running on the
GPU take longer for small numbers of combinations to fit.
\begin{figure}[!t]
  \centering
  \includegraphics[width=0.9\linewidth]{figures/TotalTimeOpencl_1000}
  \caption{The timing of the \textsc{opencl} algorithm as a function of the
    number of track fits. We compare timing running the algorithm serially on a CPU, 
    using \textsc{opencl} on the CPU, and on an AMD GPU.}
  \label{fig:opencl_timing}
\end{figure}
Surprisingly, though \textsc{opencl} is an open standard, we find that
we are unable to run the same \textsc{opencl} code on Nvidia GPUs. On two
different installations with two different video cards, we find
incorrect results running the code that ran successfully on the AMD
card. At present it is unclear what the root cause  is.
